---
title: "Hayman Fire Recovery"
author: "Sam Leuthold"
date: "1/31/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, warning=F,message=F}
library(tidyverse)
library(lubridate)
library(ggpubr)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways

knitr::opts_knit$set(root.dir='..')
```


```{r dataread, warning=F,message=F}

#Reading in files
files <- list.files('./data',full.names=T)


#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')


ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))

```




## Question 1

What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlation s a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months

```{r, warning=F,message=F, fig.height= 7.5, fig.width= 6, fig.align='center'}

## Create tibble with individual columns, filter summer months

Q1.data <- full_long %>%
           filter(!is.na(value)) %>%
           pivot_wider(names_from = c(data),
                       values_from = c(value)) %>% 
           mutate(month = month(DateTime, label = T)) %>%
           filter(month %in% c("Jun", "Jul", "Aug", "Sep")) %>%
           mutate(site = recode(site, unburned = "Unburned", burned = "Burned"))

## Create another tibble with all the data named "all".

Q1.all <- full_long %>%
           filter(!is.na(value)) %>%
           pivot_wider(names_from = c(data),
                       values_from = c(value)) %>% 
           mutate(month = month(DateTime, label = T)) %>%
           filter(month %in% c("Jun", "Jul", "Aug", "Sep")) %>%
           mutate(site = recode(site, unburned = "Unburned", burned = "Burned")) %>%
           mutate(month = "All")

## Combine datasets

Q1.data <- rbind(Q1.data, Q1.all)

###

## Create an empty tibble to fill with correlations between NDVI and NDMI for each month, and each site type.

Q1.corr <- tibble(month = rep(unique(Q1.data$month), 2),
                  site = rep(unique(Q1.data$site), each = 5),
                  Corr = vector(mode = "numeric", length = 10))

## Loop through and filter out subsets of data to run cor() on.

for(i in 1:10){

  temp <- Q1.data %>% 
            filter(month == Q1.corr$month[i] &
                     site == Q1.corr$site[i]) # It's unclear to me if this is "tidy" approved. Is there a better way to do it?
  
  Q1.corr$Corr[i] <- paste("r = ", round(cor(temp$ndmi, temp$ndvi, "complete.obs"),2))
}

###

## Plot data.

ggplot(data = Q1.data, aes(x = ndmi, y = ndvi)) +
  geom_point(fill = "lightgrey", 
             shape = 21, 
             size = 3) +
  geom_smooth(method = "lm",
              se = FALSE,
              lty = 1,
              color = "black",
              size = 1.5) +
  facet_grid(month~site) +
  scale_x_continuous(name = "NDMI", 
                     limits = c(-0.4, 0.4)) +
  scale_y_continuous(name = "NDVI", 
                     limits = c(0, 0.75)) +
  geom_text(data = Q1.corr,
            aes(x = 0.3,
                y = 0.65,
                label = Corr,
                fill = NULL,
                group = month)) +
  theme_classic2() +
  theme(panel.background = element_rect(fill = NULL, color = "black"),
        strip.background = element_blank(),
        axis.text = element_text(color = "black"))

```

## Question 2 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?

```{r, warning=F, message=F, fig.height= 6, fig.width= 6, fig.align='center'}

## Create a dataset with just the winter values, then find the annual average winter NDSI

Q2.data.winter <- full_long %>%
                    filter(!is.na(value)) %>%
                    pivot_wider(names_from = c(data),
                                values_from = c(value)) %>% 
                    mutate(Year = year(DateTime), 
                           Month = month(DateTime)) %>%
                    filter(Month %in% c(1, 2, 3, 4)) %>%
                    group_by(site, Year) %>%
                    summarise(Mean_Winter_NDSI = mean(ndsi)) 

## Create a dataset with just the summer values, then find the annual average summer NDVI

Q2.data.summer <- full_long %>%
                    filter(!is.na(value)) %>%
                    pivot_wider(names_from = c(data),
                                values_from = c(value)) %>% 
                    mutate(Year = year(DateTime), 
                           Month = month(DateTime)) %>%
                    filter(Month %in% c(6, 7, 8)) %>%
                    group_by(site, Year) %>%
                    summarise(Mean_Summer_NDVI = mean(ndvi)) 


## Combine the datasets

Q2.data <-  inner_join(Q2.data.summer,
                       Q2.data.winter)

## Find the correlation between summer NDVI and winter NDSI

Q2.Corr <- paste("r = ", 
                 round(cor(Q2.data$Mean_Summer_NDVI,
                           Q2.data$Mean_Winter_NDSI,
                           use = "complete.obs"), 3))

## Plot data

ggplot(data = Q2.data, aes(x = Mean_Winter_NDSI, y = Mean_Summer_NDVI)) +
  geom_point(fill = "lightgrey", 
             shape = 21, 
             size = 4) +
  geom_smooth(method = "lm",
              se = FALSE,
              lty = 2,
              color = "black",
              size = 1.5) +
  scale_x_continuous(name = "Mean Winter NDSI",
                     limits = c(-0.5, 0.6)) +
  scale_y_continuous(name = "Mean Summer NDVI",
                     limits = c(0.11, 0.6)) +
  annotate(geom = "text",
           x = 0.5,
           y = 0.6, 
           label = Q2.Corr) +
  theme_classic2() +
  theme(panel.background = element_rect(fill = NULL, color = "black"),
        axis.text = element_text(color = "black"))
```


## Question 3

How is the snow effect from question 2 different between pre- and post-burn
and burned and unburned? 

```{r, warning=F, message=F, fig.height= 6, fig.width= 6, fig.align='center'}

## Add column for site burn timing to Q2.data.

Q3.data <-  Q2.data %>% 
              mutate(Status = if_else(Year < 2002, "Pre-Burn", "Post-Burn")) %>%
              mutate(site = recode(site, unburned = "Unburned", burned = "Burned"))

## Use the same for loop structure from Q 1 to calculate correlations for each subset of data.

Q3.corr <- tibble(Status = rep(unique(Q3.data$Status), 2),
                  site = rep(unique(Q3.data$site), each = 2),
                  Corr = vector(mode = "numeric", length = 4))

## Loop through and filter out subsets of data to run cor() on.

for(i in 1:4){

  temp <- Q3.data %>% 
            filter(Status == Q3.corr$Status[i] &
                     site == Q3.corr$site[i])
  
  Q3.corr$Corr[i] <- paste("r = ", round(cor(temp$Mean_Winter_NDSI, temp$Mean_Summer_NDVI, "complete.obs"),3))
}

###

## Plot data

ggplot(data = Q3.data, aes(x = Mean_Winter_NDSI, y = Mean_Summer_NDVI)) +
  geom_point(fill = "lightgrey", 
             shape = 21, 
             size = 4) +
  geom_smooth(method = "lm",
              se = FALSE,
              lty = 2,
              color = "black",
              size = 1.5) +
  facet_grid(Status~site) +
  scale_x_continuous(name = "Mean Winter NDSI",
                     limits = c(-0.5, 0.6)) +
  scale_y_continuous(name = "Mean Summer NDVI",
                     limits = c(0.11, 0.6)) +
    geom_text(data = Q3.corr,
            aes(x = 0.45,
                y = 0.55,
                label = Corr)) +
  theme_classic2() +
  theme(panel.background = element_rect(fill = NULL, color = "black"),
        axis.text = element_text(color = "black"))
  

```

## Question 4

What month is the greenest month on average? 

```{r, warning=F, message=F, fig.height= 6, fig.width= 6, fig.align='center'}

Q4.data <- full_long %>%
            filter(!is.na(value)) %>%
            pivot_wider(names_from = c(data),
                        values_from = c(value)) %>% 
            mutate(Year = year(DateTime), 
                   Month = month(DateTime, label = T)) %>%
            group_by(Month) %>% 
            summarise(Mean_NDVI = mean(ndvi, na.rm = T))

Q4.data %>%
  filter(Mean_NDVI == max(Mean_NDVI))


ggplot(data = Q4.data, aes(x = Month, y = Mean_NDVI)) +
  geom_col(fill = "lightgrey",
           color = "black") +
  geom_col(data = Q4.data[Q4.data$Month == "Aug",], 
           aes(x = Month, y = Mean_NDVI), 
           fill = "forestgreen", 
           color = "black") +
  geom_hline(yintercept = 0) +
  ylab("Mean NDVI") +
  theme_classic2() +
  theme(panel.background = element_rect(fill = NULL, color = "black"),
        axis.text = element_text(color = "black"))

```


## Question 5

What month is the snowiest on average?

```{r, warning=F, message=F, fig.height= 6, fig.width= 6, fig.align='center'}

Q5.data <- full_long %>%
            filter(!is.na(value)) %>%
            pivot_wider(names_from = c(data),
                        values_from = c(value)) %>% 
            mutate(Year = year(DateTime), 
                   Month = month(DateTime, label = T)) %>%
            group_by(Month) %>% 
            summarise(Mean_NDSI = mean(ndsi, na.rm = T))

Q5.data %>%
  filter(Mean_NDSI == max(Mean_NDSI))


ggplot(data = Q5.data, aes(x = Month, y = Mean_NDSI)) +
  geom_col(fill = "lightgrey",
           color = "black") +
  geom_col(data = Q5.data[Q5.data$Month == "Jan",], 
           aes(x = Month, y = Mean_NDSI), 
           fill = "red", 
           color = "black") +
  geom_hline(yintercept = 0) +
  ylab("Mean NDSI") +
  theme_classic2() +
  theme(panel.background = element_rect(fill = NULL, color = "black"),
        axis.text = element_text(color = "black"))

```


## Bonus Question: Redo all problems with `spread` and `gather` using modern tidyverse syntax. 


## Bonus Question: Use Climage Engine to pull the same data for the assignment, but updated with 2020/2021 data.




